%! Author = murfe
%! Date = 08.06.2026

\chapter{Grundlagen}\label{ch:grundlagen}
\textcolor{red}{Orientierung ca. 8-12 Seiten}
%TODO: Einleitung
%TODO: auf jeden Fall irgendwo at most once / at least once / effectively exactly once aufgreifen


\section{Hintergrundverarbeitung in verteilten Systemen}\label{sec:hintergrundverarbeitung}

\subsection{Synchrone vs. Asynchrone Verarbeitung}\label{subsec:synchroneVsAsynchron}

"A consequence of queueing is also that consumers are generally \textit{asynchronous}: when a producer sends
a message, it normally only waits for the broker to confirm that it has buffered the message and
does not wait for the message to be processed by consumers. The delivery to consumers will happen at
some undetermined future point in time—often within a fraction of a second, but sometimes
significantly later if there is a queue backlog."\parencite[Kap.11 --> Message brokers]{kleppmann_designing_2017}

\subsection{Background Job-Processing}\label{subsec:backgroundJobs}

"In principle, a file or database is sufficient to connect producers and consumers: a producer writes
every event that it generates to the datastore, and each consumer periodically polls the datastore
to check for events that have appeared since it last ran. This is essentially what a batch process
does when it processes a day’s worth of data at the end of every day." \parencite[Kap.11 --> Transmitting Event Streams]{kleppmann_designing_2017}

"What happens if nodes crash or temporarily go offline—are any messages lost? As with databases,
durability may require some combination of writing to disk and/or replication (see the sidebar
\href{https://learning.oreilly.com/library/view/designing-data-intensive-applications/9781491903063/ch07.html#sidebar_transactions_durability}{“Replication and Durability”}), which has a cost. If you can afford to sometimes lose
messages, you can probably get higher throughput and lower latency on the same hardware."\parencite[Kap.11 Messaging Systems]{kleppmann_designing_2017}


\section{Konsistenz und konkurrierende Verarbeitung}\label{sec:konsistenz}
\textcite[Kap. 7 --> Transactions]{kleppmann_designing_2017}

\subsection{ACID}\label{subsec:acid}

\subsection{Locking}\label{subsec:locking}

\subsection{Transaktionsgrenzen}\label{subsec:transaktionsgrenzen}


\section{Fehlertoleranz und Wiederherstellung}\label{sec:fehlertoleranz}

\href{https://learning.oreilly.com/library/view/designing-data-intensive-applications/9781491903063/ch01.html#sec_introduction_reliability}{“Reliability”}

\subsection{Retry-/Backoff-Strategien}\label{subsec:retry/backoff}

\subsection{Fehlerklassifikationen}\label{subsec:fehlerklassifikationen}

\subsection{Recovery}\label{subsec:recovery}

Dein „remote datastore“ ist praktisch deine PostgreSQL-Jobtabelle (persistenter Job-Store). Das entspricht dem Ansatz „State remote halten“, was Recovery vereinfacht, aber Laufzeit/Contention beeinflusst.
 Dein Worker hat vermutlich lokalen, transienten Zustand während der Verarbeitung (in-memory). Der muss nach Crash nicht „rekonstruiert“ werden, sondern du brauchst Mechanismen wie Lease/Timeout + Retry, um Jobs wieder aufzunehmen.
„Replay state from input“ entspricht bei dir am ehesten: Job erneut ausführen (at-least-once) – und genau deshalb brauchst du Idempotenz, weil „replay“ bei Side-Effects sonst doppelte Ergebnisse erzeugt.
\parencite[Kap.11 Rebuilding a state failure]{kleppmann_designing_2017}

\subsection{Idempotenz}\label{subsec:idempotenz}
Wichtig für exactly-once Semantik
\parencite[Kap.11 --> Atomic commit revisited & Idempotence]{kleppmann_designing_2017}
\parencite[Kap.3.3.6]{richardson_microservices_2018}
message Handler, um doppelte Nachrichten zu Eleminieren \parencite[Kap.3.3.6]{richardson_microservices_2018}

\section{Zustandsmodellierung mittels State Machines}\label{sec:zustandsmodellierung}


\section{Stand der Technik}\label{sec:forschungsstand}
